Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-02-24 WSSSPE1_report/WSSSPE1_summary_report.tex
adding text on kitware

\subsection{Scientific Software}
Special purpose software that was developed specifically for a target use-case
scenario or keeping some fixed/frozen requirements in mind. Software projects
pertaining to specific scientific domain often tend to be in a niche and the
user community tends to be small to medium. Numerical accuracy and algorithmic
optimization are some of the paramount requirements of most scientific
software. Advantages of a smaller codebase and requirements include stability,
ease of installation, configurability and integration into larger systems.
While the software can stay stable and require relatively low maintenance, the
responsibility is often on the shoulders of a very few developers. Community
participation in the development tends to be low. Development is linear and
simplistic with a limited scope to follow software best practices. Domain
science oriented software developed to solve  specific problems, e.g.,
\emph{DUNE}, \emph{PETSc}, \emph{MINRES-QLP}, \emph{FASTMath}. One exception is
\emph{Kitware}. Sustainability of such software is often a significant
challenge. Many submissions report the software considered as a `byproduct` of
actual research. Others contend that the software was not the main funded part
of their research.

\subsection{Infrastructure Software}
Infrastructure oriented software developed to better utilize new
infrastructure, e.g., \emph{MVAPICH}, \emph{visIT}. Explores uncharted
territories. Highest risk since the projects are experimental. Overlap with the
utility software as it must serve as a facilitator to run existing software.
Long requirement-development-test cycles. Wide community support but tends to
be handled by specialists rather then end users. Promotes collaborations, one
of the key ingredient of a sustainable process.
Ferenbaugh~\cite{Ferenbaugh_WSSSPE} discusses the risks and challenges
associated with software projects dealing with new architectures.

>>>>

\subsection{Scientific Software}
Special purpose software that was developed specifically for a target use-case
scenario or keeping some fixed/frozen requirements in mind. Software projects
pertaining to specific scientific domain often tend to be in a niche and the
user community tends to be small to medium. Numerical accuracy and algorithmic
optimization are some of the paramount requirements of most scientific
software. Advantages of a smaller codebase and requirements include stability,
ease of installation, configurability and integration into larger systems.
While the software can stay stable and require relatively low maintenance, the
responsibility is often on the shoulders of a very few developers. Community
participation in the development tends to be low. Development is linear and
simplistic with a limited scope to follow software best practices. Domain
science oriented software developed to solve  specific problems, e.g.,
\emph{DUNE}, \emph{PETSc}, \emph{MINRES-QLP}, \emph{FASTMath}. Sustainability
of such software is often a significant challenge. Many submissions report the
software considered as a `byproduct` of actual research.  Others contend that
the software was not the main funded part of their research. One exception is
\emph{Kitware}. While being a software product specializing in scientific
process, developing communities around software process is a core development
focus. One example of this process is the development of the \emph{CMake} build
utility which started out as a building tool for \emph{ITK} but grown to become
a generic build utility for C++ projects.

\subsection{Infrastructure Software}
Software developed to better utilize new infrastructure, e.g., \emph{MVAPICH},
\emph{visIT}. Explores uncharted territories. Highest risk since the projects
are experimental. Successful projects reap high rewards and have longer usage
span. By nature, they overlap with the utility software as it must serve as a
facilitator to run existing software. Long requirement-development-test cycles.
Wide community support but tends to be handled by specialists rather then end
users. Promotes collaborations across the breadth (eg. different departments)
and depth (eg. stakeholders within a department) of community, one of the key
ingredient of a sustainable process. Ferenbaugh~\cite{Ferenbaugh_WSSSPE}
discusses the risks and challenges associated with software projects dealing
with new architectures.
